Natural selection in the Health and Retirement Study

### 2023-09-17 hrs-selection.qmd
Use withReplicates for main bootstraps

# TODO: think about your bootstraps, what's the test... given that 
# the scores are not independent, does it really make sense to aggregate
# a general size comparison? Or am I secretly just reproducing my own
# crappy standard errors?

@@

Appendix @fig-fertility shows effect sizes of PGS correlations with realised fertility among each ethnic group. Standard errors are large because of the relatively low sample sizes, and no scores are significant at Bonferroni-corrected p \< 0.05/`r n_tests`. But I am most concerned with looking at patterns across scores rather than judging the significance of individual scores.

@@

n_reps <- 599
cors_boot <- list()
for (eth in eths) {
  rand_short <- rand |> 
    subset(ethnicity == eth)
  cors_boot[[eth]] <-
    replicate(n_reps, {
      # survey package automatically does the right thing
      # possibly I should be using survey::bootweights
      # or survey::withReplicates and a function
      r_boot <- rand_short[sample(nrow(rand_short), replace = TRUE), ]

      coef_pgs_fert <- pgs |> 
        map(\(x) reformulate(c(x, pcs), response = "raevbrn")) |> 
        map(\(x) svyglm(x, r_boot)) |> 
        map_dbl(\(x) coef(x)[[2]])

      coef_pgs_sibs <- pgs |> 
        map(\(x) reformulate(c(x, pcs), response = "r10sibs")) |> 
        # need to add , weights = weight/(r10sibs + 1)
        map(\(x) svyglm(x, r_boot)) |> 
        map_dbl(\(x) coef(x)[[2]])

      coef_pgs_edyrs <- pgs |> 
        map(\(x) reformulate(c(x, pcs), response = "raedyrs")) |> 
        map(\(x) svyglm(x, r_boot)) |> 
        map_dbl(\(x) coef(x)[[2]])

      c(fert_edyrs = cor(coef_pgs_fert, coef_pgs_edyrs),
        sibs_edyrs = cor(coef_pgs_sibs, coef_pgs_edyrs))
    })
}

tails_boot_fert <- map(cors_boot, 
                       \(x) quantile(x["fert_edyrs", ], c(0.025, 0.975)))

tails_boot_sibs <- map(cors_boot,
                       \(x) quantile(x["sibs_edyrs", ], c(0.025, 0.975)))

if (! all(tails_boot_fert$white < 0)) {
  stop("Fertility-edu correlation NOT sig at 95% for whites")
}
if (all(tails_boot_fert$black < 0)) {
  stop("Fertility-edu correlation IS sig at 95% for blacks")
}

@@

rand |> 
    mutate(
      Ethnicity = fct_relevel(titles[ethnicity], "White", "Black")
    ) |> 
    ggplot(aes(rabyear, fill = Ethnicity)) + 
      geom_bar(position = "identity") +
      labs(x = "Birth year", y = "Count") +
      scale_fill_manual(values = c(Black = "orange2", White = "steelblue4")) +
      theme_minimal() +
      theme(legend.title = element_blank())

>>>>

# TODO: think about your bootstraps, what's the test... given that 
# the scores are not independent, does it really make sense to aggregate
# a general size comparison? Or am I secretly just reproducing my own
# crappy standard errors?
# 
# TODO: is ADHD reverse-coded? What about XDISORDER?

@@

```
Appendix @fig-fertility shows effect sizes of PGS correlations with realised fertility among each ethnic group. Standard errors are large because of the relatively low sample sizes, and no scores are significant at Bonferroni-corrected p \< 0.05/`r n_tests`. But I am most concerned with looking at patterns across scores rather than judging the significance of individual scores.

@@

n_reps <- 599

calc_cors_edu <- function (weights, data, var) {
  coef_pgs_fert <- pgs |> 
        map(\(x) reformulate(c(x, pcs), response = var)) |> 
        map(\(x) lm(x, data, weights = weights)) |> 
        map_dbl(\(x) coef(x)[[2]])

  coef_pgs_edu <- pgs |> 
        map(\(x) reformulate(c(x, pcs), response = "raedyrs")) |> 
        map(\(x) lm(x, data, weights = weights)) |> 
        map_dbl(\(x) coef(x)[[2]])

  cor(coef_pgs_edu, coef_pgs_fert)
}

cors_fert_edu <- cors_sibs_edu <- list()
for (eth in eths) {
  rand_boot <- rand |> 
    subset(ethnicity == eth) |> 
    as.svrepdesign(type = "bootstrap", replicates = n_reps)
  r_parents_boot <- r_parents |> 
    subset(ethnicity == eth) |> 
    as.svrepdesign(type = "bootstrap", replicates = n_reps)

  cors_fert_edu[[eth]] <- 
    withReplicates(rand_boot, theta = calc_cors_edu, var = "raevbrn")
  cors_sibs_edu[[eth]] <- 
    withReplicates(r_parents_boot, theta = calc_cors_edu,var = "r10sibs")
}

tails_boot_fert <- map(cors_fert_edu, 
                       \(x) x + c(-1.96, 1.96) * SE(x))

tails_boot_sibs <- map(cors_sibs_edu, 
                       \(x) x + c(-1.96, 1.96) * SE(x))
```

@@

rand$variables |> 
    mutate(
      Ethnicity = fct_relevel(titles[ethnicity], "White", "Black")
    ) |> 
    ggplot(aes(rabyear, fill = Ethnicity)) + 
      geom_bar(position = "identity") +
      labs(x = "Birth year", y = "Count") +
      scale_fill_manual(values = c(Black = "orange2", White = "steelblue4")) +
      theme_minimal() +
      theme(legend.title = element_blank())
